\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-01-25 draft.tex
remove formal lemma 1 and its proof; move statement about proofs in appendix to after result 1

\begin{lemma}
  $\tau^n(\ga(e_n,\ve_n))$ is increasing in $\ga(e_n,\ve_n)$.
  \label{res:tninc}
\end{lemma}
All proofs are in the Appendix.

@@

\noindent Lobbying affects only the weight the legislature places on the profits of the import-competing industry. These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement. However, there are decreasing returns for the lobby: increasing effort leads to less than proportional increases in the probability that the trade agreement will be broken.

@@

\section{Appendix}
\label{sec:appendix}
\noindent \textbf{\hypertarget{tninc}{Proof of Lemma~\ref{res:tninc}}}:\\
Expand the first order condition for the legislature's problem as follows:
   \begin{equation}
        - D(P_{X}) \frac{\partial P_X}{\partial \tau} + \ga(e_n,\ve_n) Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}
	    + \left[ D(P_{X})- Q_X(P_{X})\right] + \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) = 0
			\label{eq:foc}
	 \end{equation}
I will refer to this equation as \textit{FOC}.

By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = - \frac{\frac{\partial FOC}{\partial \ga}}{\frac{\partial FOC}{\partial \tau^n}}$. We have $\frac{\partial FOC}{\partial \ga} = Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}$ from the proof of Lemma 0. This is always positive. Lemma 0 also shows that $\frac{\partial FOC}{\partial \tau^n}$ is negative everywhere. Thus $\frac{\partial \tau^n}{\partial \ga}\geq 0$. $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\
Substituting from Equation~\ref{eq:ml}, Equation~\ref{eq:b} can be re-written as
\begin{multline}  
  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \mathit{PS}_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
\end{multline}
Rearranging, we have $b(e_b,\bta,\btn) = $
\begin{multline}  
  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \mathit{PS}_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \mathit{PS}_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
  \label{eq:b_ex}
\end{multline}
%\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. By Assumption~\ref{as:ga_c}, the right side of the inequality is increasing and concave in $e_b$. Thus $b(e_b,\bta,\btn)$ is increasing and concave in $e_b$.    $\hfill\blacksquare$
\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. Call it $Z$. Thus we have $b(e_b,\bta,\btn) = \Pr [Z < \ga(e_b,\ve_b) ] = 1 - F_{\ga}(\ga=Z)$ where $F_{\ga}(\ga=Z) = F_{\ta}(\ta=h^{-1}(\ga,e))$ by the Change of Variables Theorem and Assumption~\ref{as:ga_ta} with $\ga = h(e,\ta)$ giving the change of variable.

>>>>

@@

\noindent All proofs are in the Appendix.

Lobbying affects only the weight the legislature places on the profits of the import-competing industry. These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement. However, there are decreasing returns for the lobby: increasing effort leads to less than proportional increases in the probability that the trade agreement will be broken.

@@

\section{Appendix}
\label{sec:appendix}
\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\
Substituting from Equation~\ref{eq:ml}, Equation~\ref{eq:b} can be re-written as
\begin{multline}  
  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \mathit{PS}_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
\end{multline}
Rearranging, we have $b(e_b,\bta,\btn) = $
\begin{multline}  
  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \mathit{PS}_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \mathit{PS}_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
  \label{eq:b_ex}
\end{multline}
%\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. By Assumption~\ref{as:ga_c}, the right side of the inequality is increasing and concave in $e_b$. Thus $b(e_b,\bta,\btn)$ is increasing and concave in $e_b$.    $\hfill\blacksquare$
\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. Call it $Z$. Thus we have $b(e_b,\bta,\btn) = \Pr [Z < \ga(e_b,\ve_b) ] = 1 - F_{\ga}(\ga=Z)$ where $F_{\ga}(\ga=Z) = F_{\ta}(\ta=h^{-1}(\ga,e))$ by the Change of Variables Theorem and Assumption~\ref{as:ga_ta} with $\ga = h(e,\ta)$ giving the change of variable.
